---
title: "Female reproductive tract module validation report" # Title of the document
subtitle: "Levonorgestrel simulations" # Subtitle of the document. May be left empty
author: "" # Author of the document. May be left empty
institute: "ESQlabs GmbH"
date: "now"
date-format: "YYYY-MM-DD HH:mm:ss" 
params:  
  loadPreSimulatedResults: FALSE # If `TRUE`, results will be loaded from the folder specified in `loadResultsFolder`
  loadResultsFolder: "Levonorgestrel" # Sub-folder where simulated results is located, in case `loadPreSimulatedResults` is `TRUE`. The folder must be located in the "SimulationResults" folder defined in the `ProjectConfiguration`.
output-file: "Female reproductive tract module validation report - Levonorgestrel" #Name of the output file.
link-citations: TRUE
reference-section-title: "References"
bibliography: references_levonorgestrel.bib
execute:
  echo: false #do not show code chunks
  output: true
  warning: false # do not print warning
  cache: false # do not use cache unless specified in chunks
format:
  pdf:
    toc: true # display table of content (TOC)
    toc-depth: 3 # display three levels in TOC
    toc-title: Table of Content # override default "Content" TOC title
    lof: false # Print list of figures?
    lot: false # Print list of tables?
    fig-cap-location: bottom # figure captions location
    tbl-cap-location: top # table captions location
    number-sections: true # add number to sections
    colorlinks: true # links format
    fig-dpi: 300 # plot quality
    fig-width: 6 # plot quality
    fig-asp: 0.618 # h/w ratio
    fig-pos: "H" # ensure plots don't move to bottom of doc
    fig-align: center # plots are centered by default
    papersize: A4
   # fontfamily: "Segoe UI" # Can we use "Segoe UI Light" here?
---

```{r initialize, include = FALSE, eval = TRUE}
# Load required libraries
library(esqlabsR)
# Source all utility files
sourceAll("utils")
# Create project configuration
projectConfiguration <- esqlabsR::createProjectConfiguration(path = "../ProjectConfiguration.xlsx")

## Set name of the observed data file
projectConfiguration$dataFile <- "Levonorgestrel_TimeValuesData.xlsx"
```

```{r simulateScenarios1, include = FALSE, eval = TRUE}
# This code runs the defined scenarios, generates figures, etc. This chunk can be re-used multiple times
# in the report. When re-using, remember to re-name the chunk as every chunk must have a unique name.

# Define the scenarios to execute, the sheets to load data from, and the plots to generate in this run
scenarioNames <- list(
  "FRT_LNG_oral_singleDose",
  "FRT_LNG_oral_multipleDose", 
  "Mirena_WeibullRelease"
)

dataSheets <- c(
  "LNG"
)
  
plotGridNames <- c(
  "Mirena_WeibullRelease", 
  "LNG_plasma_single",
  "LNG_Nilsson_oral"
)

outputFolder <- "Levonorgestrel"

# Adjust simulation run options, if necessary.
# E.g. disable check for negative values if required
simulationRunOptions <- NULL
# simulationRunOptions <- ospsuite::SimulationRunOptions$new()
# simulationRunOptions$checkForNegativeValues <- FALSE

reportResults <- createResults(
  projectConfiguration = projectConfiguration,
  params = params,
  scenarioNames = scenarioNames,
  customParams = NULL,
  dataSheets = dataSheets,
  customDataSets = dataSets,
  plotGridNames = plotGridNames,
  simulationRunOptions = simulationRunOptions
)

# A separate set of scenarios can be executed e.g. with specified custom parameters.
# The two sets of results can be conbined with the function `combineResults()`.

# Optionally - export the figures to PNG files
# exportPlotsToPNG(reportResults$plots)
```

\newpage

::: callout-note
```{r resultsOrigin, echo=FALSE, results='asis'}
if (params$loadPreSimulatedResults) {
  cat("This report has been created with simulation results loaded from results folder '", params$loadResultsFolder, "'.")
} else {
  cat("This report has been created by running the simulations. The results have been stored in folder '", reportResults$scenarioResults$outputFolder, "'.")
}
```
:::

\newpage

# Introduction

This report presents the validation of the female reproductive tract module using Levonorgestrel simulations. The module is designed to simulate the pharmacokinetics of drug targetting tissues in the female reproductive tract, and distribution of the drug upon local administration (e.g., vaginal, intrauterine or cervical).

Levonorgestrel is a synthetic progestogen widely used in various hormonal contraceptive formulations. A PBPK model for oral adminstration of levonorgestrel was obtained from literature [@cicali_quantitative_2021], and extended for intrauterine (IUD) administration.

# Compound: Levonorgestrel 1

## Parameters

Name                                             | Value         | Value Origin                                                   | Alternative          | Default
------------------------------------------------ | ------------- | -------------------------------------------------------------- | -------------------- | -------
Solubility at reference pH                       | 14 mg/l       | Other-Assumption-Biorelevant solubility estimated using Simcyp | Estimated FaSSIF     | True   
Reference pH                                     | 6.5           | Other-Assumption-Biorelevant solubility estimated using Simcyp | Estimated FaSSIF     | True   
Lipophilicity                                    | 4 Log Units   | Database-Manual Fit                                            | logP                 | True   
Fraction unbound (plasma, reference value)       | 1             | Other-Other-Binding kinetics operating in mechanistic mode     | fu                   | True   
Permeability                                     | 0.31 cm/min   |                                                                | Parameter Estimation | False  
Specific intestinal permeability (transcellular) | 0.0002 cm/min |                                                                | Fitted               | True   
Is small molecule                                | Yes           |                                                                |                      |        
Molecular weight                                 | 312.44 g/mol  |                                                                |                      |        
Plasma protein binding partner                   | Unknown       |                                                                |                      |        
Density (drug)                                   | 1.2 g/cm³     | Unknown                                                        |                      |        


## Calculation methods

Name                    | Value              
----------------------- | -------------------
Partition coefficients  | Rodgers and Rowland
Cellular permeabilities | PK-Sim Standard    


## Processes

### Systemic Process: Total Hepatic Clearance-Parameter identification

Species: Human

#### Parameters

Name                          | Value          | Value Origin                                                                                                         
----------------------------- | -------------- | ---------------------------------------------------------------------------------------------------------------------
Body weight                   | 60 kg          | Unknown                                                                                                              
Hematocrit                    | 0.41           | Unknown                                                                                                              
Fraction unbound (experiment) | 0.014          |                                                                                                                      
Lipophilicity (experiment)    | 3.65 Log Units |                                                                                                                      
Plasma clearance              | 0.1 l/h/kg     | Parameter Identification


### Metabolizing Enzyme: CYP3A4-Parameter Identification

Molecule: CYP3A4

Metabolite: Metabolite

#### Parameters

Name                           | Value                        | Value Origin            
------------------------------ | ---------------------------- | ------------------------
In vitro CL/recombinant enzyme | 1.45 µl/min/pmol rec. enzyme | Parameter Identification


### Specific Binding: SHBG-Qi-Gui & Humpel 1990

Molecule: SHBG

#### Parameters

Name | Value      | Value Origin
---- | ---------- | ------------:
koff | 17.4 1/min |             
Kd   | 0.0005 µM  |             


### Specific Binding: ALB-Bayer report

Molecule: ALB

#### Parameters

Name | Value     | Value Origin
---- | --------- | ------------:
koff | 6.5 1/min |             
Kd   | 6 µmol/l  |             


# Clinical Data

Data from the following studies has been used for model validation [@cicali_quantitative_2021; @apter_pharmacokinetics_2014; @nilsson_tissue_1982]:

| Reference             | Administration route     | Dose         | Clinical data                        | Population                | Weight (kg) | Height (cm) | Age (years) | Number of participants |
|-----------------------|--------------------------|--------------|---------------------------------------|------------------------------|-------------|-------------|-------------|------------------------|
| Cicali 2021       | PO              | 0.09 mg     | Plasma | European ICRP (female)       |    -    |      -    |       -  | <20   |
| Nilsson 1982      | PO              | 0.25 mg       | Plasma; Fat; Endometrium; Myometrium | European ICRP (female)       |   -     |     -    |   -     | 4  |    
| Apter 2014       | Intrauterine (IUD)              | 20 µg/24h       | Plasma | European ICRP (female)       |    -     |     -     |    -     | <15   |    

# Results

Comparison of individual simulations with observed data are presented in the following figures.

```{r, include = TRUE, eval = TRUE}
#| label: fig-Levonorgestrel-LNG_plasma_single
#| fig-cap: "Concentration-time profiles (a) and goodness-of-fit plots (b) of levonorgestrel administered as a single 0.09 mg oral dose (a, b) administration. Lines represent predicted concentration for the mean individual. Points (errorbars) represent mean (standard deviation) observed data [@cicali_quantitative_2021]. Solid line represent unity, and dashed lines represent 2-fold deviation. "

reportResults$plots$`LNG_plasma_single`
```

```{r, include = TRUE, eval = TRUE, fig.asp = 0.8}
#| label: fig-Levonorgestrel-LNG_Nilsson_oral
#| fig-cap: "Concentration-time profiles (a) and goodness-of-fit plots (b) of levonorgestrel administered as a multiple 0.25 mg oral dose (a, b) administration. Lines represent predicted concentration for the mean individual. Points (errorbars) represent mean (standard deviation) observed data [@nilsson_tissue_1982]. Solid line represent unity, and dashed lines represent 2-fold deviation."


reportResults$plots$`LNG_Nilsson_oral`
```

```{r, include = TRUE, eval = TRUE}
#| label: fig-Levonorgestrel-Mirena_WeibullRelease
#| fig-cap: "Concentration-time profiles (a) and goodness-of-fit plots (b) of levonorgestrel administered as a 52 mg intrauterine (a, b) administration. Lines represent predicted concentration for the mean individual. Points (errorbars) represent mean (standard deviation) observed data [@apter_pharmacokinetics_2014]. Solid line represent unity, and dashed lines represent 2-fold deviation. "

reportResults$plots$`Mirena_WeibullRelease`
```

# References


